Formal Grammars as Types in Non-commutative Bunched Type Theory

### 2023-12-14 paper/paper.tex
draft comments, typo fixes, and some comments

\newcommand{\cat}[1]{\mathbf{#1}}

@@

The category of grammars is quite rich in structure, which can be most
easily observed from the fact that it is equivalently defined as the
category of functors $\Set^{\Sigma^*}$ where $\Sigma^*$ is here viewed
as a discrete category, that is, the objects are strings and only
morphisms are identity morphisms. Such functor categories, often
called presheaf categories, are incredibly rich in structure, which we
will exploit in section \ref{blah} to succinctly define the semantics
of regular and context-free grammars.

@@

The proposed axioms are a direct translation of the Kleene algebra
axioms to a categorical setting. Its most unusual aspect is the
axiomatization of the Kleene star as a family os initial algebras
satisfying certain isomorphisms. If the Kleene category $\cat{K}$ has
more structure, then these isomorphisms hold ``for free''.

>>>>

\newcommand{\cat}[1]{\mathbf{#1}}
\newcommand{\Set}{\mathbf{Set}}

\newif\ifdraft
\drafttrue
\newcommand{\steven}[1]{\ifdraft{\color{orange}[{\bf Steven}: #1]}\fi}
\newcommand{\max}[1]{\ifdraft{\color{blue}[{\bf Max}: #1]}\fi}
\newcommand{\pedro}[1]{\ifdraft{\color{red}[{\bf Pedro}: #1]}\fi}

@@

The category of grammars is quite rich in structure, which can be most
easily observed from the fact that it is equivalently defined as the
category of functors $\Set^{\Sigma^*}$ where $\Sigma^*$ is here viewed
as a discrete category, that is, the objects are strings and only
morphisms are identity morphisms. Such functor categories, often
called presheaf categories, are incredibly rich in structure, which we
will exploit in section \ref{blah} to succinctly define the semantics
of regular and context-free grammars. \steven{Topoi model dependent type theory. This dependence is seemingly useful for also reasoning about context sensitivity, but this doesn't work for \textit{reasons}, right?}

@@

\steven{What is the ``of course'' modality? Is this ``$!$'' in linear logic?}

The proposed axioms are a direct translation of the Kleene algebra
axioms to a categorical setting. Its most unusual aspect is the
axiomatization of the Kleene star as a family of initial algebras
satisfying certain isomorphisms. If the Kleene category $\cat{K}$ has
more structure, then these isomorphisms hold ``for free''.
